Gravitational wave sources in our Galactic backyard: Predictions for BH and NS binaries in LISA

### 2021-07-01 paper/tex/paper.tex
Updates from Overleaf
eccentricity and chirp mass uncertainty
add new models and change commands to match
add commands make things consistent, update table
fix typos
add email
add corresponding author
make abstract centred (typo)
some feedback updates

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{fontawesome5}

@@

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}
\newcommand{\tom}[1]{\textcolor{ForestGreen}{[Tom: #1]}}

% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

@@

%% Physics variations shortcuts
\newcommand{\modFid}{A}
\newcommand{\modBetaLow}{B}
\newcommand{\modBetaMed}{C}
\newcommand{\modBetaHigh}{D}
\newcommand{\modCaseBB}{E}
\newcommand{\modAlphaLow}{F}
\newcommand{\modAlphaHigh}{G}
\newcommand{\modOpt}{H}
\newcommand{\modRapid}{I}
\newcommand{\modNSLow}{J}
\newcommand{\modNSHigh}{K}
\newcommand{\modNoPISN}{L}
\newcommand{\modSigLow}{M}
\newcommand{\modSigLower}{N}
\newcommand{\modNoBH}{O}

\newcommand{\modRangeMT}{B-E}
\newcommand{\modRangeCE}{F-H}
\newcommand{\modRangeSN}{I-O}

@@

@@

% selma's newest catchy one
\title{Gravitational wave sources in our Galactic backyard: LISA predictions for BH and NS binaries}

@@

\author[0000-0001-7969-1569]{S. Justham}
\affiliation{School of Astronomy \& Space Science, University of the Chinese Academy of Sciences, Beijing 100012, China}
\affiliation{\UvA}

\begin{abstract}{
    \input{sections/abstract.tex}
}
\end{abstract}

@@

\acknowledgments{We thank Mike Lau, Floris Kummer, Eva Laplace, Rob Farmer, Katie Sharpe, Mathieu Renzo, Katie Brevik, Alberto Sesana, Valeriya Korol, Will Farr and the members of the COMPAS collaboration for insightful discussions. This project was funded in part by the European Union’s Horizon 2020 research and innovation program from the European Research Council (ERC, Grant agreement No. 715063), and by the Netherlands Organization for Scientific Research (NWO) as part of the Vidi research program BinWaves with project number 639.042.728. We further acknowledge the Black Hole Initiative funded by a generous contribution of the John Templeton Foundation and the Gordon and Betty Moore Foundation.}

>>>>

\usepackage{lipsum}
\usepackage{physics}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{fontawesome5}
\usepackage{xcolor} % to add more color options

@@

\newcommand{\todo}[1]{{\color{red}{[TODO: #1}]}}
\newcommand{\tom}[1]{\textcolor{ForestGreen}{[Tom: #1]}}
\newcommand{\floor}[1]{\textcolor{Bittersweet}{[Floor: #1]}}
% custom function for adding units
\makeatletter
\newcommand{\unit}[1]{%
    \,\mathrm{#1}\checknextarg}
\newcommand{\checknextarg}{\@ifnextchar\bgroup{\gobblenextarg}{}}
\newcommand{\gobblenextarg}[1]{\,\mathrm{#1}\@ifnextchar\bgroup{\gobblenextarg}{}}
\makeatother

@@

%% Physics variations shortcuts
\newcommand{\modFid}{A}
\newcommand{\modBetaLow}{B}
\newcommand{\modBetaMed}{C}
\newcommand{\modBetaHigh}{D}
\newcommand{\modCaseBB}{E}
\newcommand{\modCaseBBOpt}{F}
\newcommand{\modAlphaLowest}{G}
\newcommand{\modAlphaLow}{H}
\newcommand{\modAlphaHigh}{I}
\newcommand{\modAlphaHighest}{J}
\newcommand{\modOpt}{K}
\newcommand{\modRapid}{L}
\newcommand{\modNSLow}{M}
\newcommand{\modNSHigh}{N}
\newcommand{\modNoPISN}{O}
\newcommand{\modSigLow}{P}
\newcommand{\modSigLower}{Q}
\newcommand{\modNoBH}{R}
\newcommand{\modWRLow}{S}
\newcommand{\modWRHigh}{T}

\newcommand{\modRangeMT}{B-F}
\newcommand{\modRangeCE}{G-K}
\newcommand{\modRangeSN}{L-R}
\newcommand{\modRangeML}{S-T}

\newcommand{\nModels}{20}
\newcommand{\nMinusOneModels}{19}

\newcommand{\BHBHFourYear}{36}
\newcommand{\BHNSFourYear}{33}
\newcommand{\NSNSFourYear}{8}
\newcommand{\BHBHTenYear}{56}
\newcommand{\BHNSTenYear}{55}
\newcommand{\NSNSTenYear}{13}

\newcommand{\BHBHNotWDWDFour}{15}
\newcommand{\BHNSNotWDWDFour}{10}
\newcommand{\NSNSNotWDWDFour}{4}
\newcommand{\BHBHNotWDWDTen}{22}
\newcommand{\BHNSNotWDWDTen}{17}
\newcommand{\NSNSNotWDWDTen}{7}

\newcommand{\BHBHDistinguishedFour}{5}
\newcommand{\BHNSDistinguishedFour}{4}
\newcommand{\NSNSDistinguishedFour}{3}
\newcommand{\BHBHDistinguishedTen}{11}
\newcommand{\BHNSDistinguishedTen}{7}
\newcommand{\NSNSDistinguishedTen}{5}

@@

% \color{white}
% \pagecolor{black}

@@

% selma's newest catchy one
\title{Gravitational wave sources in our Galactic backyard: Predictions for BH and NS binaries in LISA}

@@

\author[0000-0001-7969-1569]{S. Justham}
\affiliation{School of Astronomy \& Space Science, University of the Chinese Academy of Sciences, Beijing 100012, China}
\affiliation{\UvA}
\affiliation{\mpa}

\correspondingauthor{Tom Wagg}
\email{tomjwagg@gmail.com}

\begin{abstract}
    \input{sections/abstract.tex}
\end{abstract}

@@

\acknowledgments{We thank Mike Lau, Floris Kummer, Eva Laplace, Rob Farmer, Katie Sharpe, Mathieu Renzo, Katie Brevik, Alberto Sesana, Valeriya Korol, Will Farr, Tyson Littenberg and the members of the COMPAS collaboration for insightful discussions. This project was funded in part by the European Union’s Horizon 2020 research and innovation program from the European Research Council (ERC, Grant agreement No. 715063), and by the Netherlands Organization for Scientific Research (NWO) as part of the Vidi research program BinWaves with project number 639.042.728. We further acknowledge the Black Hole Initiative funded by a generous contribution of the John Templeton Foundation and the Gordon and Betty Moore Foundation.}

### 2021-07-01 paper/tex/sections/appendix.tex
finish distinguishing
update appendix table
fix page positions in appendix

\section{Detection Rate Normalisation}\label{app:rate_normalisation}
In this section we explain in detail the normalisation process that we refer to in Section~\ref{sec:gw_detection}. From each instance of the Milky Way that we simulate we can extract the fraction of targets that are detectable, where we define a target as one of BHBH, BHNS or NSNS that merges in a Hubble time. Note that we complete this calculation separately for each DCO type and each model variations.

We now explain how to convert the detectable fraction to a detection rate for the Milky Way. We write that the \textit{number} of detectable targets in the Milky Way is
\begin{equation}\label{eq:norm_frame}
    N_{\rm detect} = f_{\rm detect} \cdot N_{\rm target, MW},
\end{equation}
where $f_{\rm detect}$ is the fraction of targets in the instance that were detectable and $N_{\rm target, MW}$ is the total number of targets that have been formed in the Milky Way's history. We can further break this total down into
\begin{equation}
    N_{\rm target, MW} = \avg{ \mathcal{R}_{\rm target} } \cdot M_{\rm SF, MW},
\end{equation}
where $\avg{ \mathcal{R}_{\rm target} }$ is the average number of targets formed per star forming mass and $M_{\rm SF, MW}$ is the star forming mass of the Milky Way, meaning the total mass of every star ever formed in the Milky Way. We explain how to calculate each of these terms in respective subsections below.

\subsection{Average target formation rate}
Double compact object formation is strongly metallicity dependent, so we must find the average rate as the integral over metallicity
\begin{equation}\label{eq:norm_avg_target_formation}
    \avg{ \mathcal{R}_{\rm target} } = \int_{Z_{\rm min}}^{Z_{\rm max}} p_{Z} \mathcal{R}_{\rm target, Z} \dd{Z},
\end{equation}
where $Z_{\rm min}, Z_{\rm max}$ are the minimum and maximum sampled metallicities, $p_Z$ is the probability of forming a star at the metallicity $Z$ (which can be found using the distribution in \citealp{Frankel+2018}) and $\mathcal{R}_{\rm target, Z}$ is the number of targets formed per star forming mass.
\begin{equation}
    \mathcal{R}_{\rm target, Z} =  \frac{N_{\rm target, Z}}{ M_{\rm SF, Z} }.
\end{equation}
In reality, this integral is instead approximated as a sum over the metallicity bins that we use in our simulation. The number of targets in our sample at a metallicity $Z$, $N_{\rm target, Z}$, can be written simply as the sum of the targets' weights:
\begin{equation}
    N_{\rm target, Z} = \sum_{i=1}^{N_{\rm binaries, Z}} w_i \theta_{\rm target, i},
\end{equation}
where $w_i$ is the adaptive importance sampling weight assigned to the binary by STROOPWAFEL, $N_{\rm binaries, Z}$ is the number of binaries at metallicity $Z$ in our sample and $\theta_{\rm target, i}$ is a step function that is only $1$ when the binary is a target and otherwise $0$.

@@

\citet{Licquia+2015} find that the total stellar mass today in the Milky Way is $6.08 \pm 1.14 \times 10^{10} \unit{M_{\odot}}$. This total includes all stars and stellar remnants (white dwarfs, neutrons stars and black holes) but \textit{excludes} brown dwarfs. We can write that the total mass of every star every formed in the Milky Way is
\begin{equation}\label{eq:m_SF_MW}
    M_{\rm SF, MW} = 6.08 \pm 1.14 \times 10^{10} \unit{M_{\rm \odot}} \cdot \frac{\avg{m}_{\rm SF, total}}{\avg{m}_{\rm SF, today}},
\end{equation}
where $\avg{m}_{\rm SF, total}$ is the average mass of a star over the history of the Milky Way and is defined as
\begin{equation}
    \avg{m}_{\rm SF, total} = \int_{0}^{t_{\rm MW}} p_{\rm birth}(\tau) \int_{0.01}^{200} \zeta(m)\ m \dd{m} \dd{\tau},
\end{equation}
where $t_{\rm MW}$ is the age of the Milky Way, $\zeta(m)$ is the \citet{Kroupa+2001} IMF function and $p_{\rm birth}(\tau)$ is the probability of a star being formed at a lookback time $\tau$ (Eq.~\ref{eq:thin_disc_tau}). $\avg{m}_{\rm SF, today}$ is the average mass of all stars and stellar remnants (excluding brown dwarfs) present in the Milky Way today is defined as follows (note that we integrate from $0.08$ not $0.01$ since observations of today's Milky Way mass exclude brown dwarfs)
\begin{equation}
    \avg{m}_{\rm SF, today} = \int_{0}^{t_{\rm MW}} p_{\rm birth}(\tau) \int_{0.08}^{200} \zeta(m)\ m_{\rm today}(m, \avg{Z}_\tau, \tau) \dd{m} \dd{\tau},
\end{equation}
where $m_{\rm today}(m, Z, \tau)$ is the current mass of a star that was formed $\tau$ years ago at a metallicity $Z$. We calculate $m_{\rm today}(m, Z, \tau)$ by interpolating the final masses given by COMPAS for a grid of single stars over different masses and metallicities using the \citet{Fryer+2012} delayed prescription and default wind mass loss settings. We only use the final mass if more than the main sequence lifetime of the star has passed, otherwise we just use the initial mass. Since we don't know the exact metallicity of each star we use the average star forming metallicity in the Milky Way at a lookback time $\tau$ using the birth radius distribution and metallicity equation from \citet{Frankel+2018}. If we now evaluate Equation \ref{eq:m_SF_MW}, we find that the total mass of every star that has ever formed in the Milky Way is
\begin{equation}
    M_{\rm SF, MW} = 6.1 \pm 1.1 \times 10^{10} \unit{M_{\odot}} \cdot \frac{0.384 \unit{M_{\odot}}}{0.257 \unit{M_{\odot}}} = 9.1 \pm 1.1 \times 10^{10} \unit{M_{\odot}},
\end{equation}
an increase of approximately 50\% from the value still in stars today!

@@

@@

\subsection{Detection rate table}
\begin{table*}[htb]
    \centering
    \caption{The number of detectable binaries in a 4 and 10 year LISA mission for the 15 different model variations and each DCO type. Each value shows the median and the 90\% confidence interval.}
    \begin{tabular}{c|lll|lll}
        \hline
        \multirow{2}{*}{Model} & \multicolumn{3}{c|}{LISA (4 year)} & \multicolumn{3}{c}{LISA (10 year)} \\ \cline{2-7}
         & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} \\
        \hline
        A & \confinv{34.4}{20.9}{46.6} & \confinv{29.9}{20.1}{47.4} & \confinv{7.7}{5.0}{10.3} & \confinv{58.6}{31.0}{53.3} & \confinv{51.9}{28.0}{95.0} & \confinv{13.0}{6.8}{14.2}\\
        B & \confinv{33.9}{19.7}{36.5} & \confinv{15.5}{9.6}{19.6} & \confinv{2.8}{1.9}{4.3} & \confinv{55.6}{27.7}{45.9} & \confinv{26.7}{13.8}{23.7} & \confinv{4.8}{2.7}{5.6}\\
        C & \confinv{20.6}{11.1}{20.5} & \confinv{5.3}{3.0}{5.3} & \confinv{4.3}{3.1}{5.2} & \confinv{33.7}{15.7}{26.7} & \confinv{8.9}{4.2}{7.0} & \confinv{7.3}{4.5}{6.1}\\
        D & \confinv{21.9}{14.7}{37.8} & \confinv{4.1}{2.6}{4.2} & \confinv{11.6}{8.3}{22.7} & \confinv{36.0}{21.6}{45.4} & \confinv{6.8}{3.4}{5.3} & \confinv{20.5}{12.7}{27.7}\\
        E & \confinv{33.7}{19.4}{37.0} & \confinv{3.3}{2.3}{6.3} & \confinv{0.2}{0.1}{0.1} & \confinv{54.9}{26.7}{44.2} & \confinv{5.8}{3.5}{8.2} & \confinv{0.3}{0.1}{0.2}\\
        F & \confinv{26.2}{16.0}{35.8} & \confinv{15.0}{11.4}{84.0} & \confinv{5.7}{3.8}{7.3} & \confinv{43.3}{22.2}{43.7} & \confinv{27.2}{18.2}{86.3} & \confinv{9.5}{5.2}{10.0}\\
        G & \confinv{29.5}{15.9}{27.6} & \confinv{23.5}{14.0}{27.2} & \confinv{23.4}{15.2}{27.9} & \confinv{47.5}{21.4}{38.3} & \confinv{40.9}{20.7}{36.3} & \confinv{39.8}{21.4}{36.3}\\
        H & \confinv{115.1}{98.0}{123.0} & \confinv{30.7}{21.4}{163.7} & \confinv{7.4}{4.8}{14.6} & \confinv{173.6}{117.7}{150.5} & \confinv{55.4}{32.8}{185.8} & \confinv{13.0}{6.8}{25.5}\\
        I & \confinv{22.7}{12.4}{26.8} & \confinv{48.7}{30.9}{97.1} & \confinv{6.3}{4.0}{9.0} & \confinv{36.1}{16.8}{30.8} & \confinv{84.0}{45.0}{142.8} & \confinv{10.9}{5.8}{11.1}\\
        J & \confinv{47.5}{31.6}{71.9} & \confinv{24.3}{16.3}{37.4} & \confinv{6.8}{4.4}{10.9} & \confinv{80.7}{46.8}{86.8} & \confinv{41.7}{22.9}{47.0} & \confinv{11.6}{6.2}{14.0}\\
        K & \confinv{30.1}{18.0}{40.4} & \confinv{37.6}{24.6}{44.9} & \confinv{7.8}{5.1}{11.6} & \confinv{52.5}{28.2}{48.4} & \confinv{65.1}{35.0}{54.7} & \confinv{13.4}{7.1}{14.8}\\
        L & \confinv{35.2}{21.3}{45.2} & \confinv{30.2}{19.9}{49.6} & \confinv{7.9}{5.0}{11.7} & \confinv{58.9}{31.0}{54.9} & \confinv{52.4}{28.5}{96.9} & \confinv{13.3}{7.1}{14.8}\\
        M & \confinv{43.0}{28.3}{58.3} & \confinv{68.7}{43.4}{108.7} & \confinv{16.1}{10.4}{20.5} & \confinv{69.4}{38.1}{69.1} & \confinv{117.4}{61.8}{154.5} & \confinv{27.3}{14.5}{26.4}\\
        N & \confinv{53.0}{34.1}{58.3} & \confinv{110.6}{71.3}{193.1} & \confinv{35.7}{23.6}{54.8} & \confinv{84.4}{44.9}{73.3} & \confinv{194.9}{102.6}{256.0} & \confinv{60.2}{32.7}{67.8}\\
        O & \confinv{43.1}{27.0}{39.0} & \confinv{76.0}{43.9}{91.2} & \confinv{9.3}{6.0}{10.4} & \confinv{68.5}{35.8}{47.5} & \confinv{130.8}{62.9}{122.0} & \confinv{15.8}{8.4}{13.8}\\
        P & \confinv{44.3}{28.4}{62.2} & \confinv{26.7}{17.0}{48.9} & \confinv{9.0}{5.6}{10.9} & \confinv{69.9}{38.0}{78.9} & \confinv{45.9}{25.9}{64.4} & \confinv{15.3}{7.4}{13.9}\\
        Q & \confinv{0.0}{0.0}{0.0} & \confinv{6.3}{3.7}{6.1} & \confinv{7.8}{4.7}{8.0} & \confinv{0.0}{0.0}{0.0} & \confinv{10.8}{5.3}{7.6} & \confinv{13.0}{6.4}{10.4}\\
        \hline
    \end{tabular}
    \label{tab:detection_rates}
\end{table*}

>>>>

\section{Detection Rate Normalisation}\label{app:rate_normalisation}
In this section we explain the normalisation process that we refer to in Section~\ref{sec:gw_detection}. From each simulated instance of the Milky Way we can extract the fraction of targets that are detectable, where we define a target as one of BHBH, BHNS or NSNS that merges in a Hubble time.

To convert the detectable fraction to a detection rate for the Milky Way, we write that the \textit{number} of detectable targets in the Milky Way is
\begin{equation}\label{eq:norm_frame}
    N_{\rm detect} = f_{\rm detect} \cdot N_{\rm target, MW},
\end{equation}
where $f_{\rm detect}$ is the fraction of targets in the instance that were detectable and $N_{\rm target, MW}$ is the total number of targets that have been formed in the Milky Way's history. We can further break this total down into
\begin{equation}
    N_{\rm target, MW} = \avg{ \mathcal{R}_{\rm target} } \cdot M_{\rm SF, MW},
\end{equation}
where $\avg{ \mathcal{R}_{\rm target} }$ is the average number of targets formed per star forming mass and $M_{\rm SF, MW}$ is the star forming mass of the Milky Way, meaning the total mass of every star ever formed in the Milky Way.

\subsection{Average target formation rate}
Double compact object formation is metallicity dependent, so we find the average rate as the integral over metallicity, which is given by
\begin{equation}\label{eq:norm_avg_target_formation}
    \avg{ \mathcal{R}_{\rm target} } = \int_{Z_{\rm min}}^{Z_{\rm max}} p_{Z} \mathcal{R}_{\rm target, Z} \dd{Z},
\end{equation}
where $Z_{\rm min}, Z_{\rm max}$ are the minimum and maximum sampled metallicities, $p_Z$ is the probability of forming a star at the metallicity $Z$ (which can be found using the distribution in \citealp{Frankel+2018}) and $\mathcal{R}_{\rm target, Z}$ is the number of targets formed per star forming mass,
\begin{equation}
    \mathcal{R}_{\rm target, Z} =  \frac{N_{\rm target, Z}}{ M_{\rm SF, Z} }.
\end{equation}
In practice, this integral is instead approximated as a sum over the metallicity bins that we use in our simulation. The number of targets in our sample at a metallicity $Z$, $N_{\rm target, Z}$, can be written simply as the sum of the targets' weights:
\begin{equation}
    N_{\rm target, Z} = \sum_{i=1}^{N_{\rm binaries, Z}} w_i \theta_{\rm target, i},
\end{equation}
where $w_i$ is the adaptive importance sampling weight assigned to the binary by STROOPWAFEL, $N_{\rm binaries, Z}$ is the number of binaries at metallicity $Z$ in our sample and $\theta_{\rm target, i}$ is a step function that is only $1$ when the binary is a target and otherwise $0$.

@@

\citet{Licquia+2015} find that the total stellar mass today in the Milky Way is $6.08 \pm 1.14 \times 10^{10} \unit{M_{\odot}}$. This total includes all stars and stellar remnants (white dwarfs, neutrons stars and black holes) but \textit{excludes} brown dwarfs. We can write that the total mass of every star every formed in the Milky Way is
\begin{equation}\label{eq:m_SF_MW}
    M_{\rm SF, MW} = 6.08 \pm 1.14 \times 10^{10} \unit{M_{\rm \odot}} \cdot \frac{\avg{m}_{\rm SF, total}}{\avg{m}_{\rm SF, today}},
\end{equation}
where $\avg{m}_{\rm SF, total}$ is the average mass of a star over the history of the Milky Way and is defined as
\begin{equation}
    \avg{m}_{\rm SF, total} = \int_{0}^{t_{\rm MW}} p_{\rm birth}(\tau) \int_{0.01}^{200} \zeta(m)\ m \dd{m} \dd{\tau},
\end{equation}
where $t_{\rm MW}$ is the age of the Milky Way, $\zeta(m)$ is the \citet{Kroupa+2001} IMF function and $p_{\rm birth}(\tau)$ is the probability of a star being formed at a lookback time $\tau$ (Eq.~\ref{eq:thin_disc_tau}). $\avg{m}_{\rm SF, today}$ is the average mass of all stars and stellar remnants (excluding brown dwarfs) present in the Milky Way today is defined as follows (note that we integrate from $0.08$ not $0.01$ since observations of today's Milky Way mass exclude brown dwarfs)
\begin{equation}
    \avg{m}_{\rm SF, today} = \int_{0}^{t_{\rm MW}} p_{\rm birth}(\tau) \int_{0.08}^{200} \zeta(m)\ m_{\rm today}(m, \avg{Z}_\tau, \tau) \dd{m} \dd{\tau},
\end{equation}
where $m_{\rm today}(m, Z, \tau)$ is the current mass of a star that was formed $\tau$ years ago at a metallicity $Z$. We calculate $m_{\rm today}(m, Z, \tau)$ by interpolating the final masses given by COMPAS for a grid of single stars over different masses and metallicities using the \citet{Fryer+2012} delayed prescription and default wind mass loss settings. We only use the final mass if more than the main sequence lifetime of the star has passed, otherwise we just use the initial mass. Since we don't know the exact metallicity of each star we use the average star forming metallicity in the Milky Way at a lookback time $\tau$ using the birth radius distribution and metallicity equation from \citet{Frankel+2018}. If we now evaluate Equation \ref{eq:m_SF_MW}, we find that the total mass of every star that has ever formed in the Milky Way is
\begin{equation}
    M_{\rm SF, MW} = (6.1 \pm 1.1) \times 10^{10} \unit{M_{\odot}} \cdot \frac{0.384 \unit{M_{\odot}}}{0.257 \unit{M_{\odot}}} = (9.1 \pm 1.1) \times 10^{10} \unit{M_{\odot}},
\end{equation}
an increase of approximately 50\% from the value still in stars today!

@@

\newpage

@@

\subsection{Detection rate table}
\begin{table*}[htb]
    \centering
    \caption{The number of detectable binaries in a 4- and 10-year LISA mission for the 15 different model variations and each DCO type. Each value shows the median and the 90\% confidence interval.}
    \begin{tabular}{c|lll|lll}
        \hline
        \multirow{2}{*}{Model} & \multicolumn{3}{c|}{LISA (4 year)} & \multicolumn{3}{c}{LISA (10 year)} \\ \cline{2-7}
         & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} & \scriptsize{BHBH} & \scriptsize{BHNS} & \scriptsize{NSNS} \\
        \hline
        A & \confinv{39.6}{6.6}{6.4} & \confinv{36.5}{6.5}{6.5} & \confinv{8.7}{2.7}{3.3} & \confinv{62.9}{7.9}{8.1} & \confinv{60.7}{7.7}{7.3} & \confinv{14.4}{3.4}{3.6}\\
        B & \confinv{37.1}{6.1}{5.9} & \confinv{17.4}{4.4}{4.6} & \confinv{3.3}{2.3}{1.7} & \confinv{58.7}{7.7}{7.3} & \confinv{28.7}{5.7}{5.3} & \confinv{5.3}{2.3}{2.7}\\
        C & \confinv{22.4}{4.4}{4.6} & \confinv{5.8}{2.8}{2.2} & \confinv{4.7}{1.7}{2.3} & \confinv{35.8}{5.8}{6.2} & \confinv{9.4}{3.4}{2.6} & \confinv{7.6}{2.6}{2.4}\\
        D & \confinv{26.2}{5.2}{4.8} & \confinv{4.3}{2.3}{1.7} & \confinv{14.1}{4.1}{3.9} & \confinv{40.5}{6.5}{6.5} & \confinv{7.1}{3.1}{2.9} & \confinv{23.2}{5.2}{4.8}\\
        E & \confinv{36.9}{5.9}{6.1} & \confinv{4.2}{2.2}{1.8} & \confinv{0.2}{0.2}{0.8} & \confinv{58.6}{7.6}{7.4} & \confinv{6.6}{2.6}{2.4} & \confinv{0.3}{0.3}{0.7}\\
        F & \confinv{30.2}{5.2}{5.8} & \confinv{22.9}{4.9}{5.1} & \confinv{6.4}{2.4}{2.6} & \confinv{47.5}{6.5}{6.5} & \confinv{35.5}{5.5}{5.5} & \confinv{10.4}{3.4}{3.6}\\
        G & \confinv{31.8}{5.8}{5.2} & \confinv{25.9}{4.9}{5.1} & \confinv{25.9}{4.9}{5.1} & \confinv{50.5}{7.5}{7.5} & \confinv{43.5}{6.5}{6.5} & \confinv{42.4}{6.4}{6.6}\\
        H & \confinv{117.8}{10.8}{11.2} & \confinv{51.7}{6.7}{7.3} & \confinv{10.1}{3.1}{2.9} & \confinv{181.6}{13.6}{13.4} & \confinv{87.7}{9.7}{9.3} & \confinv{16.7}{3.7}{4.3}\\
        I & \confinv{25.5}{5.5}{5.5} & \confinv{62.0}{8.0}{8.0} & \confinv{7.2}{2.2}{2.8} & \confinv{39.1}{6.1}{5.9} & \confinv{102.3}{10.3}{9.7} & \confinv{11.9}{3.9}{3.1}\\
        J & \confinv{55.5}{7.5}{7.5} & \confinv{27.9}{4.9}{5.1} & \confinv{8.0}{3.0}{3.0} & \confinv{87.8}{9.8}{9.2} & \confinv{45.7}{6.7}{6.3} & \confinv{13.0}{4.0}{4.0}\\
        K & \confinv{34.8}{5.8}{6.2} & \confinv{42.4}{6.4}{6.6} & \confinv{9.0}{3.0}{3.0} & \confinv{56.0}{8.0}{7.0} & \confinv{69.5}{8.5}{8.5} & \confinv{14.8}{3.8}{4.2}\\
        L & \confinv{39.8}{6.8}{6.2} & \confinv{36.7}{5.7}{6.3} & \confinv{9.1}{3.1}{2.9} & \confinv{63.4}{8.4}{7.6} & \confinv{61.9}{7.9}{8.1} & \confinv{14.7}{3.7}{3.3}\\
        M & \confinv{48.2}{7.2}{6.8} & \confinv{80.8}{8.8}{9.2} & \confinv{18.0}{4.0}{4.0} & \confinv{75.3}{8.3}{8.7} & \confinv{134.5}{11.5}{11.5} & \confinv{29.5}{5.5}{5.5}\\
        N & \confinv{57.5}{7.5}{7.5} & \confinv{138.2}{12.2}{11.8} & \confinv{41.9}{6.9}{6.1} & \confinv{89.5}{9.5}{9.5} & \confinv{227.5}{15.5}{15.5} & \confinv{67.4}{8.4}{8.6}\\
        O & \confinv{45.7}{6.7}{6.3} & \confinv{85.3}{9.3}{9.7} & \confinv{10.2}{3.2}{2.8} & \confinv{70.9}{8.9}{8.1} & \confinv{141.6}{11.6}{12.4} & \confinv{16.8}{3.8}{4.2}\\
        P & \confinv{50.8}{6.8}{7.2} & \confinv{32.7}{5.7}{5.3} & \confinv{10.0}{3.0}{3.0} & \confinv{77.6}{8.6}{8.4} & \confinv{53.1}{7.1}{6.9} & \confinv{16.5}{4.5}{4.5}\\
        Q & \confinv{0.0}{0.0}{0.0} & \confinv{6.7}{2.7}{2.3} & \confinv{8.5}{2.5}{2.5} & \confinv{0.0}{0.0}{0.0} & \confinv{11.3}{3.3}{3.7} & \confinv{13.8}{3.8}{4.2}\\
        \hline
    \end{tabular}
    \label{tab:detection_rates}
\end{table*}

### 2021-07-01 paper/tex/sections/comparisons.tex
move figure AGAIN

In a seminal work, \citet{Nelemans+2001} were the first to investigate the population of LISA detectable stellar mass double compact objects. Their predictions of the number of detectable DCOs are very different from our own, where we find a significantly higher detection rate for BHBHs and BHNSs, as well as a slightly lower rate for NSNSs. However, since their paper, there have been many changes both to the specifications of LISA (such as the mission length and SNR threshold for detection) and our understanding of massive star evolution, which likely strongly affect the expected detections rates.

@@

The discrepancy between their results and ours could be a result of different implicit assumptions in their population synthesis code, however we expect that BSE should be similar to COMPAS since it provided the basis of the COMPAS code. Another major difference is that, unlike this work, they assume that all binaries are circular for the purpose of detection in LISA, which could result in a lower number of detections by missing eccentric binaries that appear as weaker signals when assumed to be circular. We also improve upon this work by using a larger number of metallicity bins since a low number of metallicity bins can produce artificial features in the mass distribution of DCOs and possibly affect the detection rate (Kummer et al. (in prep)).

\citet{Lau+2020} investigated the number of Galactic NSNS binaries that could be detected by LISA. Their study uses the same population synthesis code, COMPAS, as this work, though an earlier version. Despite this, their study found a much larger number of detections. They make several different physical assumptions in their population synthesis, using the \citet{Fryer+2012} \textit{rapid} remnant mass prescription, limiting the maximum neutron star mass to $2 \unit{M_{\odot}}$ and not implementing PISN. However, we note that none of these assumptions strongly affect the NSNS LISA detection rate (see bottom panel of Fig.\,\ref{fig:detection_rates}, models \modRapid{}, \modNSLow{} and \modNoPISN{}).

>>>>

\citet{Nelemans+2001} were the first to investigate the population of LISA detectable stellar mass double compact objects. Their predictions of the number of detectable DCOs are very different from our own, where we find a significantly higher detection rate for BHBHs and BHNSs, as well as a slightly lower rate for NSNSs. However, since the publication of their paper, there have been many changes both to the specifications of LISA (such as the mission length and SNR threshold for detection) and our understanding of massive star evolution, which strongly affect the expected detections rates.

@@

The discrepancy between their results and ours could be a result of different implicit assumptions in their population synthesis code, however we expect that BSE should be similar to COMPAS since it provided the basis of the COMPAS code. Another major difference is that, unlike this work, they assume that all binaries are circular for the purpose of detection in LISA, which could result in a lower number of detections by missing eccentric binaries that appear as weaker signals when assumed to be circular. We also improve upon this work by using a larger number of metallicity bins since a low number of metallicity bins can produce artificial features in the mass distribution of DCOs and possibly affect the detection rate.

\citet{Lau+2020} focussed on the number of Galactic NSNS binaries that could be detected by LISA. Their study uses the same population synthesis code, COMPAS, as this work, though an earlier version. Despite this, their study found a much larger number of detections. They make several different physical assumptions in their population synthesis, using the \citet{Fryer+2012} \textit{rapid} remnant mass prescription, limiting the maximum neutron star mass to $2 \unit{M_{\odot}}$ and not implementing PISN. However, we note that none of these assumptions strongly affect the NSNS LISA detection rate (see bottom panel of Fig.\,\ref{fig:detection_rates}, models \modRapid{}, \modNSLow{} and \modNoPISN{}).
